### 2014-11-11 paper/results_3d.tex
updated external torque estimates using goldreich

\section{Three-dimensional simulations}\label{results3d}
In this section we review 3D simulations carried 
out using ZEUS-MP and PLUTO. The main purpose is to verify 
the above results with different numerical codes, and validate  
the 2D approximation.

The 3D disc has radial size
$[r_\mathrm{min},r_\mathrm{max}]=[0.4,10]R_0$ and vertical extent  
$n_H=2$ scale-heights at $R=R_0$. The resolution is $N_r\times N_\theta\times
N_\phi=256\times32\times256$. This corresponds to about $4$ cells per
$H$in radius. Because of the much reduced resolution 
compared to 2D, we use a smooth perturbation by setting
$\delta = 10^{-3}$ and $M=1$ in Eq. \ref{randpert}. This corresponds
to a single $m=1$ disturbance in $R\in[R_1,R_2]$, which was found to dominate
the 2D simulation.

The 3D discs are initialised in approximate equilibrium only, so we
first evolve the disc without perturbations using  
$(\lmax,\mmax)=(32,0)$ up to $t=10P_0$, during which 
meridional velocities are damped out. We then restart the simulation
with the above perturbation and $(\lmax,\mmax)=(32,32)$ and damp
merdional velocities near the radial boundaries.

Fig. \ref{3d_ampmax} plots the amplitude of the $m=1$ spirals measured
in the ZEUS-MP and PLUTO runs. 
%why no high m
                                %modes as in 2D for iso disc?
                                %resolution?
Results from ZEUS-MP appears to be off-set from that obtained using
PLUTO.  
%We suspect this is due to larger numerical noise in the
%ZEUS-MP code, 
In ZEUS-MP we observe disturbances developed near
the inner boundary, which is probably responsible for the initial growth
for $t<70P_0$. This growth is likely a numerical artifact and effectively seeds
the ZEUS-UP simulation with a larger initial perturbation.  However,
once the coherent $m=1$ spiral begins to grow ($t\gtrsim 100P_0$), 
we measure similar growth rates in both codes: 
\begin{align*}
  &\gamma \simeq 0.0073\Omega_k(R_0) \quad\quad \mathrm{PLUTO},\\
  &\gamma \simeq 0.0084\Omega_k(R_0) \quad\quad \varA{ZEUS-MP}.
\end{align*}
Both are smaller than the 2D simulations. This may be due to the
low resolutions adopted and/or the small softening length employed in
2D.  Fig. \ref{3d_ampmax} also show results from strictly isothermal
simulations, which display no growth compared to cases with a 
temperature gradient.

% We also show results from strictly
% isothermal discs ($q=0$). 
%growth rate does not depend on pert amp

\begin{figure}
  \includegraphics[width=\linewidth]{figures/m1_analysis_plot_ampmax3d.ps}
  \caption{Evolution of the maximum $m=1$ density in  $r\in[R_1,R_2]$
    measured in the 3D simulations. Results from discs with a imposed 
    temperature gradient ($q=1$) and a strictly isothermal disc
    ($q=0$) are shown.  
    \label{3d_ampmax}}   
\end{figure}

Visualisations of the 3D simluations are shown in   
Fig. \ref{3d_prelim} for the disc midplane and near the upper disc
boundary. The snapshots are chosen when the spirals in the two codes
have reached similar amplitudes. The agreement between the codes,
as well as with the 2D simulations, is satisfactory.    
Away from the disc midplane, a coherent spiral arm is
not prominent, but the disturbance is radially more global, extending
up to $R\sim 4R_0$. 
%implications?

\begin{figure}
  \begin{center}
    \subfigure[ZEUS-MP]{
      \includegraphics[scale=0.382]{figures/polarxy_dens015_z0}
      \includegraphics[scale=0.382,clip=true,trim=2.24cm 0cm 0cm
      0cm]{figures/polarxy_dens015_zmax}  
    }
    \subfigure[PLUTO]{
      \includegraphics[scale=0.382]{figures/pdisk_020_z0}
      \includegraphics[scale=0.382,clip=true,trim=2.24cm 0cm 0cm
      0cm]{figures/pdisk_020_zmax}
    }
  \end{center}
  \caption{Three-dimensional simulations using the ZEUS-MP (top) and 
    PLUTO (bottom) codes. The non-axisymmetric density at the midplane (left) and
    two scale-heights above the midplane (right) is shown. Here $\psi
    \equiv \pi/2 - \theta$ is the angular height  
    from the midplane.\label{3d_prelim}}   
\end{figure}

\subsection{Vertical structure}
Fig. \ref{3d_rz} shows the vertical structure of the one-arm
spiral in the PLUTO run. The main spiral is confined to $z < H$. Thus,
a 2D disc model, representing dynamics near the disc midplane, 
should be sufficient to study the instability. Interestingly, however, for $R>2R_0$ the
spirals  increase in amplitude away from the midplane. It remains
small  in our disc mode ($|\Delta\rho| < 0.1$) l, but could become 
significant with a larger vertical domain. This shows that 
the one-arm spiral can have a global influence.

\begin{figure}
  \includegraphics[scale=0.47,clip=true,trim=0cm 0.79cm 0cm 0cm]{figures/pdisk_rz_020_sg.ps}
  \includegraphics[scale=0.47,clip=true,trim=0cm 0cm 0cm 0.64cm]{figures/pdisk_rz_020_nsg.ps}
  \caption{Non-axisymmetric density in the meridional plane from the
    PLUTO simulation. The slices are taken at the azimuth of   
    $\mathrm{max}[\Delta\rho(r,\pi/2,\phi)]$. Arrows correspond to the vector 
    $(v_r/R_0,-v_\theta/rh\sin^2{\theta})$. The top (bottom) panel corresponds
    to the inner (outer) portions of the disc. 
    \label{3d_rz}} 
\end{figure}

Although Fig. \ref{3d_rz} appears to show significant vertical motion,
the three-dimensionality parameter $\Theta < 10^{-2}$
(Eq. \ref{theta}), implying the kinetic energy associated with
vertical motion is insignificant compared to that in horizontal
motion. We find  $|v_z/c_s|\lesssim
0.2$ for $h=0.05$. Thus, vertical motions are sub-sonic.  
%This again supports the 2D approximation, altthou

%Vertical velocities
%do increase with the growth of the one-arm spiral, 
%Larger amplitudes
%are reached in ZEUS because of larger numerical noise. 
%but we find  $\Theta \lesssim 10^{-2}$.

%\begin{figure}
%  \includegraphics[width=\linewidth]{figures/m1_analysis_plot_vert.ps}
% \caption{Evolution of three-dimensionality $\Theta$, as defined by
%    Eq. \ref{theta}, which corresponds to the
%   average vertical kinetic energy density, scaled by 
%    the average horizontal kinetic energy.\label{3d_vert}} 
%\end{figure}

\subsection{Angular momentum evolution} 
Fig. \ref{3d_angmom} shows the angular momentum evolution in the 3D
runs. These plots are again similar to the 2D runs, i.e. angular
momentum lost by $J_1$ is gained by $J_0$. ZEUS-MP does not conserve
angular momentum very well, but the variation $|\Delta J/J|<
O(10^{-6})$ is small compared 
to the individual components $|\Delta J_{0,1}/J|\sim 
5\times10^{-5}$. The PLUTO run reaches similar values of
$|J_{0,1}|$, but achieves better conervation, with $|\Delta
J/J|=O(10^{-8})$.

\begin{figure}
%scale=0.41
  \includegraphics[scale=.41,clip=true,trim=0cm 1cm 0cm 0cm]{figures/nonaxi_evol_ang_zeus}
  \includegraphics[scale=.41]{figures/nonaxi_evol_ang_pluto}
  \caption{Evolution of angular momentum components in the 3D 
    simulations. The perturbation 
    relative to $t=10P_0$, during the growth of the one-arm spiral,
    is shown in units of the initial total angular momentum
    $J_\mathrm{ref}$.\label{3d_angmom}}  
\end{figure}