### 2015-01-01 paper/paper.tex

\documentclass[useAMS,usenatbib]{mn2e}
\usepackage{epsfig,rotate,graphicx}
\usepackage[fleqn]{amsmath}
\usepackage{subfigure}
\usepackage{lscape}
\usepackage{bm}

\newcommand{\p}{\partial}
\newcommand{\mnras}{MNRAS}
\newcommand{\apj}{ApJ}
\newcommand{\aj}{AJ}
\newcommand{\aap}{A\&A}
\newcommand{\aapr}{A\&ARv}
\newcommand{\apss}{ApSS}
\newcommand{\apjl}{ApJL}
\newcommand{\pasj}{PASJ}
\newcommand{\araa}{ARAA}
\newcommand{\dd}{\delta}
\newcommand{\adot}{\dot{a}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\gtrsim}{\;\raisebox{-.8ex}{$\buildrel{\textstyle>}\over\sim$}\;}
\newcommand{\lesssim}{\; \raisebox{-.8ex}{$\buildrel{\textstyle<}\over\sim$}\;}
\newcommand{\anrev}{{\it ARA\&A, }}
\newcommand{\apjs}{{\it ApJS, }}
\newcommand{\icar}{{\it Icarus, }}
\newcommand{\mnr}{{\it MNRAS, }}
\newcommand{\nat}{{\it Nature, }}
\newcommand{\sci}{{\it Sci, }}
\newcommand{\ana}{{\it A\&A, }}
\newcommand{\anas}{{\it A\&AS, }}
\newcommand{\aaps}{{\it A\&AS, }}
\newcommand{\anar}{{\it A\&AR, }}
\newcommand{\prd}{{\it Phys. Rev. D }}
\newcommand{\qjl}{{\it QJRAS, }}
\newcommand{\sbar}{\bar{\sigma}}
\newcommand{\avg}[1]{\langle #1 \rangle}
\newcommand{\bu}{\bm{u}}
\newcommand{\lmax}{l_\mathrm{max}}
\newcommand{\mmax}{m_\mathrm{max}}
\newcommand{\trel}{t_\mathrm{rel}}
\newcommand{\zeus}{{\tt ZEUS-MP }}
\newcommand{\ii}{\mathrm{i}}
\newcommand{\bv}{\bm{v}}
\newcommand{\rin}{r_\mathrm{in}}
\newcommand{\rout}{r_\mathrm{out}}
\newcommand{\ciso}{c_\mathrm{iso}}
\newcommand{\tbeta}{\tilde{\beta}}
\newcommand{\teta}{\tilde{\eta}}
\newcommand{\tcool}{t_\mathrm{cool}}
\newcommand{\jlin}{j_\mathrm{lin}}
\newcommand{\jlintot}{J_\mathrm{lin}}
\newcommand{\tbg}{T_\mathrm{BG}}
\newcommand{\varA}[1]{{\operatorname{#1}}}
\DeclareMathOperator{\erf}{erf}
\DeclareMathOperator{\real}{Re}
\DeclareMathOperator{\imag}{Im}

\usepackage{array,booktabs,tabularx}
\newcolumntype{R}{>{\centering\arraybackslash}X} % right justified tabularx columns

\title[One-armed spirals]{One-armed spirals in 
  locally isothermal, radially structured self-gravitating discs}

\author[Lin]{Min-Kai Lin$^{1}$
  \thanks{ minkailin@email.arizona.edu} \\ 
  $^1$Department of Astronomy and Steward Observatory, University of
  Arizona, 933 North Cherry Avenue, Tucson, AZ 85721, USA 
}

\begin{document}

\maketitle
\begin{abstract} 
  {\bf We describe a new mechanism that leads to the destabilisation of
  one-armed spiral or eccentric modes in astrophysical discs with 
  an imposed radial temperature gradient.  We use linear density wave
  theory 
  to show that non-axisymmetric perturbations generally do not conserve
  their angular momentum in the presence of a forced temperature
  gradient. This implies an exchange of angular momentum
  between linear perturbations and the background disc. In particular, 
  when the disturbance is a low-frequency, one-armed
  trailing spiral and the disc temperature decreases outwards, this 
  interaction is unstable and leads to the growth of the spiral mode.  
  We demonstrate this phenomenon through numerical hydrodynamic
  simulations of locally isothermal discs in 2D using the FARGO
  code and in 3D with  the ZEUS-MP and PLUTO codes.  We consider
  radially structured discs with a self-gravitating sub-portion which remains
  stable in the absence of a temperature gradient. However,
  when a temperature gradient is imposed we observe exponential growth of an one-armed  
  spiral mode (azimuthal wavenumber $m=1$) with co-rotation radius 
  outside the bulk of the spiral arm, resulting in a low-frequency  
  spiral pattern in the self-gravitating region of the disc.  The development of this one-armed 
 spiral does not require the movement of the central star, as found in previous studies. 
 Because destabilisation by the imposed temperature gradient does 
 not explicitly depend on self-gravity, we suggest this mechanism may also
 affect low-frequency one-armed oscillations in non-self-gravitating
 discs. 
}
\end{abstract}

%This might represent, for example, mass  
 % built-up in a circumstellar or protoplanetary
 % disc at the outer edge of a `dead zone', in which there
  %is inefficient magneto-hydrodynamic turbulence to enable accretion.

\begin{keywords}
  accretion, accretion discs, hydrodynamics, instabilities, methods: numerical, protoplanetary discs 
\end{keywords}

\section{Introduction}\label{intro}
% Protoplanetary discs are expected to contain complex structures
% \citep{armitage11}, which directly impact planet formation theories.  
An exciting development in the study of circumstellar 
discs is the direct observation of large-scale, non-axisymmetric
structures within them. These
include lopsided dust distributions 
\citep{marel13,fukagawa13,casassus13,isella13,perez14,follette14} and
spiral arms 
\citep{hashimoto11,muto12,boccaletti14,grady13,christiaens14,avenhaus14}.

% planet is popular answer for spiral structure
The attractive explanation for asymmetries in circumstellar discs is
disc-planet interaction. In particular, spiral structures  
naturally arise from the gravitational interaction between a
planet and the gaseous protoplanetary disc it is embedded in
\citep[see, e.g.][for a recent review]{baruteau13b}. Thus, the
presence of spiral arms in circumstellar discs could be signposts of 
planet formation.

% but also natural in GI.
However, spiral arms are also associated with global gravitational 
instability (GI) of differentially rotating discs
\citep{goldreich65,laughlin96b,laughlin98,nelson98,lodato05,forgan11}. Large-scale
spiral arms can provide significant angular momentum transport
necessary for mass accretion \citep{lynden-bell72, papaloizou91,balbus99,lodato04}, and
spiral structures due to GI are potentially observable with the Atacama Large 
Millimeter/sub-millimeter Array \citep{cossins10,dipierror14}.

% when/where does it happen? formation stage/dead zones (special for ppd)
GI can be expected in the earliest stages of 
circumstellar disc formation
\citep{kratter10b,inutsuka10,tsukamoto13}, and may be possible in the
outer parts of the disc \citep{rafikov05,matzner05,kimura12}.  
% outer parts  
In the case of protoplanetary discs, which have complex structure
\citep{armitage10}, GI can also occur in `dead
zones' --- sub-portions of the disc where the magneto-rotational
instability is ineffective for angular momentum transport
\citep{gammie96,turner08,landry13} and therefore has a reduced mass
accretion rate. Dead zones may then become
self-gravitating with sufficient mass built-up
\citep{armitage01,martin12,martin12b,zhu09,zhu10,zhu10b,bae13,bae14}.

% original motivation for this work (DZGI) 
A wide range of GIs can occur in  gaseous 
astrophysical discs 
\citep{papaloizou89,christo92,christo93,hadley11,hadley14}. It is then
of theoretical importance to clarify what type of GI is relevant to 
partially self-gravitating discs, such as a protoplanetary disc containing a 
massive dead zone.  %This was indeed the original goal of this project. 
This work presents a first step towards this goal. Specifically, here we 
study a class of seemingly simple disc models which were found to 
consistently produce \emph{one-armed} spiral structures, despite not 
satisfying the usual requirements to generate them.

%through this work we identify a new method to generate one arm
%spirals 
%We therefore dedicate this study to verify their origin.

%brief history of m=1 
%stellar wobble from indiana group  
Single-arm spirals, or eccentric modes, corresponding to perturbations 
with azimuthal wavenumber $m=1$, have received interest in
the context of protoplanetary discs because of their global nature
\citep{adams89,heemskerk92,laughlin96,tremaine01,papaloizou02,hopkins10}. 
In the `SLING' mechanism proposed by \cite{shu90}, the $m=1$ spiral
instability is attributed to motion of the central star induced by the
one-armed perturbation and requires a massive disc \citep[the former may
have observable consequences, ][]{michael10}. However, our numerical
models do not have these features.

%these are self-gravitating instabilities

The aforementioned analytical studies have adopted barotropic, smooth disc models. 
In this work, we identify a mechanism to generate one-armed
spiral patterns that is related to a non-barotropic equation of
state and disc structure. We find that if the disc temperature or
sound-speed is a prescribed function of position, then 
one-armed spiral modes can grow inside a cavity that results from the
adopted disc structure.

Our study is based on direct hydrodynamic simulations using three
different grid-based codes, both in two-dimensions (2D) and
three-dimensions (3D). We also apply linear
theory to interpret simulation results. We find that, when the disc
temperature is prescribed, the usual statement for the conservation of
angular momentum for linear perturbations acquires a source term
proportional to the temperature gradient, which may act to destabilise
$m=1$ perturbations. Although we work with self-gravitating discs,
this source term does not depend explicitly on disc self-gravity. Thus, the
destabilisation effect we describe should also be applicable to
non-self-gravitating discs.

This paper is organised as follows. In \S\ref{model} we list the
governing equations and describe the disc equilibria under
consideration. We briefly review basic results from linear theory in
\S\ref{wkb} to aid our interpretation of numerical simulations. 
\S\ref{methods} describes the hydrodynamic codes
employed in this work and several diagnostic measures. Our main results are presented in 
\S\ref{results2d} for 2D simulations and \S\ref{results3d} for 3D
simulations, and we further analyse them in \S\ref{discussions}. 
We summarise in \S\ref{summary}, with a discussion on
speculations for future work.

\input{model}
\input{results}
\input{results_3d}
\input{discussion}
\input{summary}

%\section*{Acknowledgments}

\bibliographystyle{mn2e}
\bibliography{ref}

\appendix
\input{appendix}

\end{document}