### 2015-01-07 paper/discussion.tex

%to discuss - fixed central star
%application to ppd
%do smooth disc results carry over to structured disks?

\section{Discussion}\label{discussions} 
   % specific to our numerical 
   % simulations%  and may or may not be relevant to generic low-frequency
   %   % modes destabilised by a forced temperature gradient.  
   %

{\bf
  \subsection{Applicability to circumstellar discs}

  \subsubsection{Thermodynamic requirements}

  % talk about additional simulations at high Q

  %issues with fragmentation 
 % requirement for fast cooling  
  % fast cooling fragments 
  % stabilised by irradiation
  % background torque and fragmentation may not be mutually exclusive
  % (speculations) 
  % these studies do no account for possible global disk structure 
  %Clarke (2009), Rafikov (2005), and Rice & Armitage (2009)
  %no fragmentation in our case because disc is maintained at slightly
  %higher values of Q. no fragmentation of one arm spiral because
  %co-rotation radius isn't on it 
  %if irradiation 
  %one-armed low frequency spiral probably will not fragment because
  %co-rotation is far away

  A locally isothermal equation of state represents the limit
  of infinitely short cooling (and heating) timescales, so the
  disc temperature instantly returns to its initial value when
  perturbed.  A short cooling timescale $t_c$ can occur in the outer
  parts of protoplanetary discs 
  \citep{rafikov05,clarke09,rice09,cossins10}.  %rafikov07 (convection)
  However, if a disc with 
  $Q\simeq 1$ is cooled (towards zero temperature) on a timescale
  $t_c\ll\Omega_k^{-1}$, it will fragment
  \citep{gammie01,paarkedooper12}.

  %internal heating?
  To avoid fragmentation, the disc must be heated to
  maintain $Q$ above the threshold for fragmentation \citep[about
  $Q\lesssim 1.4$ for isothermal discs][]{durisen07}. This may be
  possible in the outer parts of protoplanetary discs due to heating
  by irradiation \citep{rafikov07,rice11}.

  % Numerical simulations performed by
  % \cite{rice11} indicate external irradiation can stablise
  % the disc against fragmentation.

  %It is important to note, however,
  %that these studies do not consider the possible effect of global
  %structures present in our numerical simulations.

  %chiang and goldreich
  The background torque may therefore exist in the outer
  parts of protoplanetary discs that are strongly irradiated, such
  that the disc temperature is set externally with a non-zero radial
  gradient.  Of course, if external irradiation sets a strictly
  isothermal outer disc, then the background torque vanishes.

%where does the structure come from? (mass built up, planet gap)

  \subsubsection{Radial disc structure}
  In our numerical simulations the 
  $m=1$ spiral is confined between two $Q$-barriers, where real solutions to the local
  dispersion relation is possible (Eq. \ref{wavenumber}). The
  existence of such a cavity is a result of the adopted disc
  structure. Thus, in our disc models the main role of disc structure
  and self-gravity is to 
  allow a local $m=1$ mode to be set up, which is then 
  destabilised by the background torque density.

  In order to confine an $m=1$ spiral between two $Q$-barriers, we
  should have $Q^2(1-\nu^2)=1$ at two radii. Assuming 
  Keplerian rotation and a slow pattern speed $\Omega_p\ll\Omega$, this amounts to
  \begin{align}\label{qb_cond}
    \left(\frac{R_{Qb}}{R_c}\right)^{-3/2} = 2Q^2(R_{Qb}). 
  \end{align}
  Then it may be possible to have two $Q$-barriers when the $Q$ profile
  of the disc rises more rapidly (decays more slowly) than $R^{-3/2}$
  for decreasing (increasing) $R$.

  %in fall/mass loading? 
  The surface density bump in our initial condition might represent,
  for example, mass built-up in a 
`dead zone' in a protoplanetary
  disc, where the magneto-rotational 
  instability is ineffective for angular momentum transport
  \citep{gammie96,turner08,landry13} and therefore has a reduced mass
  accretion rate. The dead zone may become 
  self-gravitating with sufficient mass built-up
  \citep{armitage01,martin12,martin12b,zhu09,zhu10,zhu10b,bae13,bae14}. 
  However, the thermodynamic conditions in a dead zone 
  %internal heating and cooling can balance such that a global radial
  %temperature gradient is maintained - then background torque may
  %exist.

  %particularly simple functional representation of the bump. a
  %realistic bump may be different

  Another possibility is a gap opened by a giant planet. In that case $Q$
  rises rapidly towards the gap centre since it is a region of low
  surface density, which may satisfy Eq. \ref{qb_cond}. Then the inner
  edge of our bump
  function mimics the outer gap edge. The outer gap edge is then a
  possible site for the growth of a low-frequency one-armed spiral
  through the background torque. 
 %or gap opened by a fragment

}

% The simulations themselves consider an isothermal or locally isothermal equation of state only.  Gammie (2001) and other work since have shown that the evolution of self-gravitating discs depends both on the stability criterion and on the balance between heating and cooling.  If the cooling rate is fast, it is likely to fragment, otherwise it will settle into a quasi-steady, self-gravitating state.

% When considering isothermal simulations, the likely evolution is one of two things.  It's either initially so unstable that the disc quickly fragments, or it's initially stable enough that any initial perturbations simply drives the system towards being even more stable and the instability quickly weakens through rapid mass transfer (see the work by Laughlin & Bodenheimer 1994/1996).  There is no real way to drive a long-lived quasi-steady, non-fragmenting, self-gravitating state because this is only possible if the system can remain quasi-stable though balancing heating and cooling (see, for example, Lodato & Rice 2004).

% In fact, in the conclusions it is suggested that the conditions used in the simulation would require a cooling < 0.1 \Omega^{-1} which, based on the work of Gammie 2001, would quickly lead to fragmentation.  The reason it doesn't in the work considered here is that the isothermal assumption assumes that there is some equivalently fast heating mechanism but, in the work here, this is not provided by the instability itself.  If it were, the simulations would become very quickly unstable.  Also, if one looks at the work of Clarke (2009), Rafikov (2005), and Rice & Armitage (2009) the only region of a realistic disc where such rapid cooling is possible is in the outer disc, but such regions are either very gravitationally unstable and fragment, or quickly stablised by an external heating source.  I guess it is possible that there could be a scenario where this exactly produces the conditions assumed here, but that would seem rather fine-tuned and it would be more likely that an external heating source would simply stabilise the disc (unless the disc were extremely massive and could remain unstable even at reasonably high disc temperatures).

{\bf
  \subsection{Motion of the central star}

  \subsection{Role of Lindblad and co-rotation torques}
  One possible effect of self-gravity is to allow the one-armed spiral  
  to act as an external potential for the exterior disc. This is
  analogous to disc-satellite interaction 
  \citep{goldreich79}, where the embedded satellite exerts a torque on
  the disc at

In Appendix \ref{disc-planet} we estimate the magnitude of this effect
  using basic results from disc-planet theory \citep[see, e.g.][and
  references therein]{papaloizou07}.

One possible effect of self-gravity is to allow the one-armed spiral  
to act as an external potential for the wider disc. This is
analogous to disc-satellite interaction 
\citep{goldreich79}. Here, we estimate the magnitude of this effect
using basic results from disc-planet theory \citep[see, e.g.][and
references therein]{papaloizou07}.

}

%Furthermore, this
%non-local interaction may be important for the long-term evolution of
%the outer disc, since it gains angular momentum from the

%also transport ang mom outwards

% Having confirmed the $m=1$ spiral instability through numerical
% simulations, we now use results from the simulations combined with
% linear theory to construct an explanation for their growth. For
% simplicity, we return to 2D discs.

% The dominant disturbance in our simulations is an one-armed, $m=1$
% trailing spiral, and its growth is associated with the imposed temperature
% gradient. As discussed in \S\ref{global_cons}, in a locally isothermal disc this imposed
% temperature gradient generally results in a torque being exerted on
% linear perturbations. An instability is therefore possible if the sign
% of this torque is the same as that of the angular momentum associated
% with the linear perturbation.

%associated angular mometum density is the same as that
%of this
% According to previous studies \citep{adams89}, our disc models
% should not support $m=1$ linear spiral instabilities. This is because
% our discs are not sufficiently massive ($M_d\lesssim 0.1M_*$) and, more
% importantly, we have surpressed the motion of the central star. The 
% latter can induce $m=1$ instabilities in massive discs  \citep{shu90}.

%the latter assumes  k=pi\Sigma G /c_s^2 is always the most favoured

% The temperature gradient affects the growth rate through the
% explicit depdence of the background torque on $q$. The growth 
% rate is also dependent on the underlying disturbance being
% considered (i.e. through $k$ and $\bm{\xi}$ in
% Eq. \ref{theoretical_rate0}). % Of course, the temperature gradient
% % plays no role on local modes. However, other disc parameters such as
% % $c_s^2$ may affect the mode that develops. 
% The dependence on $h$ in 
% Eq. \ref{theoretical_rate} assumes $k=\pi\Sigma G/c_s^2$ is always the
% most favourable wavenumber for low-frequency modes.

%By 
%definition, the global temperature gradient does not affect modes in
%local theory. 
%however, changing $h$ may also modify the mode properties (since
%temperature appears in local theory) in addition
%to the magnitude of the temperature gradient. Thus, we expect the depedence of
%the background torque on $q$ to be better described by
%Eq. \{theoretical_rate}

%by definition, q does not affect local neutral modes
%Of course, we assumed a neutral mode to obtain expressions
%for $\jlin$ and $T_\mathrm{BG}$ in the first place, 
%Recall that in local theory non-axisymmetric modes are stable for real
%$k$  
%Eq. \ref{theoretical_rate} should be interpreted
%as an order-of-magnitude estimate of the perturbative effect of a
%temperature gradient on local, low-frequency $m=1$ modes.

%neutral modes are not actually neutral)
%gamma << omega_p << omega 
%so
%Eq. \ref{theoretical_rate} is only an order-of-magnitude
%constraint. Nevertheless, the expected growth rate is in reasonable
%agreement with numerical results.

%also total torque is integral -> may get cancelations