### 2015-01-09 paper/paper.tex

\documentclass[useAMS,usenatbib]{mn2e}
\usepackage{epsfig,rotate,graphicx}
\usepackage[fleqn]{amsmath}
\usepackage{subfigure}
\usepackage{lscape}
\usepackage{bm}

\newcommand{\p}{\partial}
\newcommand{\mnras}{MNRAS}
\newcommand{\apj}{ApJ}
\newcommand{\aj}{AJ}
\newcommand{\aap}{A\&A}
\newcommand{\aapr}{A\&ARv}
\newcommand{\apss}{ApSS}
\newcommand{\apjl}{ApJL}
\newcommand{\pasj}{PASJ}
\newcommand{\araa}{ARAA}
\newcommand{\dd}{\delta}
\newcommand{\adot}{\dot{a}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\gtrsim}{\;\raisebox{-.8ex}{$\buildrel{\textstyle>}\over\sim$}\;}
\newcommand{\lesssim}{\; \raisebox{-.8ex}{$\buildrel{\textstyle<}\over\sim$}\;}
\newcommand{\anrev}{{\it ARA\&A, }}
\newcommand{\apjs}{{\it ApJS, }}
\newcommand{\icar}{{\it Icarus, }}
\newcommand{\mnr}{{\it MNRAS, }}
\newcommand{\nat}{{\it Nature, }}
\newcommand{\sci}{{\it Sci, }}
\newcommand{\ana}{{\it A\&A, }}
\newcommand{\anas}{{\it A\&AS, }}
\newcommand{\aaps}{{\it A\&AS, }}
\newcommand{\anar}{{\it A\&AR, }}
\newcommand{\prd}{{\it Phys. Rev. D }}
\newcommand{\qjl}{{\it QJRAS, }}
\newcommand{\sbar}{\bar{\sigma}}
\newcommand{\avg}[1]{\langle #1 \rangle}
\newcommand{\bu}{\bm{u}}
\newcommand{\lmax}{l_\mathrm{max}}
\newcommand{\mmax}{m_\mathrm{max}}
\newcommand{\trel}{t_\mathrm{rel}}
\newcommand{\zeus}{{\tt ZEUS-MP }}
\newcommand{\ii}{\mathrm{i}}
\newcommand{\bv}{\bm{v}}
\newcommand{\rin}{r_\mathrm{in}}
\newcommand{\rout}{r_\mathrm{out}}
\newcommand{\ciso}{c_\mathrm{iso}}
\newcommand{\tbeta}{\tilde{\beta}}
\newcommand{\teta}{\tilde{\eta}}
\newcommand{\tcool}{t_\mathrm{cool}}
\newcommand{\jlin}{j_\mathrm{lin}}
\newcommand{\jlintot}{J_\mathrm{lin}}
\newcommand{\tbg}{T_\mathrm{BG}}
\newcommand{\varA}[1]{{\operatorname{#1}}}
\DeclareMathOperator{\erf}{erf}
\DeclareMathOperator{\real}{Re}
\DeclareMathOperator{\imag}{Im}

\usepackage{array,booktabs,tabularx}
\newcolumntype{R}{>{\centering\arraybackslash}X} % right justified tabularx columns

\title[One-armed spirals]{One-armed spirals in 
  locally isothermal, radially structured self-gravitating discs}

\author[Lin]{Min-Kai Lin$^{1}$
  \thanks{ minkailin@email.arizona.edu} \\ 
  $^1$Department of Astronomy and Steward Observatory, University of
  Arizona, 933 North Cherry Avenue, Tucson, AZ 85721, USA 
}

\begin{document}

\maketitle
\begin{abstract} 
  {\bf We describe a new mechanism that leads to the destabilisation of
 non-axisymmetric waves in astrophysical discs with 
 an imposed radial temperature gradient. This might apply, for
 example, to the outer parts of protoplanetary discs. We use linear
 density wave theory 
 to show that non-axisymmetric perturbations generally do not conserve
 their angular momentum in the presence of a forced temperature
 gradient. This implies an exchange of angular momentum
 between linear perturbations and the background disc. In particular, 
 when the disturbance is a low-frequency
 trailing wave and the disc temperature decreases outwards, this 
 interaction is unstable and leads to the growth of the wave.  
 We demonstrate this phenomenon through numerical hydrodynamic
 simulations of locally isothermal discs in 2D using the FARGO
 code and in 3D with  the ZEUS-MP and PLUTO codes.  We consider
 radially structured discs with a self-gravitating region which remains
 stable in the absence of a temperature gradient. However,
 when a temperature gradient is imposed we observe exponential growth of an one-armed  
 spiral mode (azimuthal wavenumber $m=1$) with co-rotation radius 
 outside the bulk of the spiral arm, resulting in a nearly-stationary
 one-armed spiral pattern.  The development of this one-armed 
 spiral does not require the movement of the central star, as found in previous studies. 
 Because destabilisation by the imposed temperature gradient does 
 not explicitly require disc self-gravity, we suggest this mechanism may also
 affect low-frequency one-armed oscillations in non-self-gravitating
 discs. 
}
\end{abstract}

%This might represent, for example, mass  
 % built-up in a circumstellar or protoplanetary
 % disc at the outer edge of a `dead zone', in which there
  %is inefficient magneto-hydrodynamic turbulence to enable accretion.

\begin{keywords}
  accretion, accretion discs, hydrodynamics, instabilities, methods: numerical, protoplanetary discs 
\end{keywords}

\section{Introduction}\label{intro}
% Protoplanetary discs are expected to contain complex structures
% \citep{armitage11}, which directly impact planet formation theories.  
An exciting development in the study of circumstellar 
discs is the direct observation of large-scale, non-axisymmetric
structures within them. These
include lopsided dust distributions 
\citep{marel13,fukagawa13,casassus13,isella13,perez14,follette14} and
spiral arms 
\citep{hashimoto11,muto12,boccaletti14,grady13,christiaens14,avenhaus14}.

% planet is popular answer for spiral structure
The attractive explanation for asymmetries in circumstellar discs is
disc-planet interaction. In particular, spiral structures  
naturally arise from the gravitational interaction between a
planet and the gaseous protoplanetary disc it is embedded in
\citep[see, e.g.][for a recent review]{baruteau13b}. Thus, the
presence of spiral arms in circumstellar discs could be signposts of 
planet formation.

% but also natural in GI.
Spiral arms are also {\bf characteristic of} global gravitational 
instability (GI) {\bf in} differentially rotating discs
\citep{goldreich65,laughlin96b,laughlin98,nelson98,lodato05,forgan11}. Large-scale 
spiral arms can provide significant angular momentum transport
necessary for mass accretion \citep{lynden-bell72,
  papaloizou91,balbus99,lodato04}, and spiral structures due to GI are
potentially observable with the Atacama Large
Millimeter/sub-millimeter Array \citep{cossins10,dipierror14}. GI can
be expected in the earliest stages of circumstellar disc formation 
\citep{kratter10b,inutsuka10,tsukamoto13}, and may be possible in the
outer parts of the disc \citep{rafikov05,matzner05,kimura12}.

Single-arm spirals, or eccentric modes, corresponding to perturbations 
with azimuthal wavenumber $m=1$, have received interest in
the context of protoplanetary discs because of their global nature
\citep{adams89,heemskerk92,laughlin96,tremaine01,papaloizou02,hopkins10}. 
In the `SLING' mechanism proposed by \cite{shu90}, the $m=1$ {\bf
  gravitational spiral} instability is attributed to motion of the
central star induced by the one-armed perturbation and requires a
massive disc \citep[the former may have observable consequences,
][]{michael10}.

% The aforementioned analytical studies have adopted
% barotropic, smooth disc models.
% In this work, we identify a mechanism to generate one-
% armed spiral patterns that is related to a non-barotropic
% equation of state and disc structure.
% We find that if the disc temperature or sound-speed is a
% prescribed function of position, then one-armed spiral modes
% can grow inside a cavity that results from the adopted disc
% structure.

%The aforementioned analytical studies have adopted barotropic, smooth
%disc models.

{\bf
  In this work we identify a new mechanism that leads to the growth of
  one-armed spirals in astrophysical discs. We show
  that when the disc temperature is prescribed (called locally isothermal
  discs), the usual statement for the conservation of angular momentum
  for linear perturbations acquires a source term proportional to the
  temperature gradient. This permits angular momenum exchange between
  linear perturbations and the background disc. This `background
  torque' may act to destabilise low-frequency non-axisymmetric
  trailing waves when the disc temperature decrease outwards.       
}

{\bf We employ direct hydrodynamic simulations using three different
  grid-based codes to demonstrate how this `background torque' 
  can lead to the growth of one-armed spirals in radially structured, 
  self-gravitating discs. This is despite the fact that our disc
  models do not meet the requirements for the `SLING'
  mechanism. Although our numerical simulations consider
  self-gravitating discs, this `background torque' is generic for
  locally isothermal discs and its existence does not depend
  explicitly on disc self-gravity. Thus, the destabilisation effect we
  describe should also be applicable to non-self-gravitating discs.  
}

This paper is organised as follows. We describe the system of interest
and list the governing equations in 
\S\ref{model}. {\bf In \S\ref{wkb} we use linear theory  to show
  how a fixed temperature profile 
  can destabilise non-axisymmetric waves in discs.} \S\ref{methods} describes
the {\bf numerical setup and} hydrodynamic codes {\bf we use to 
  demonstrate the growth of one-armed spirals due to an imposed
  temperature gradient.} Our {\bf
  simulation} results are presented in  \S\ref{results2d} for two-
dimensional (2D) {\bf
  discs} and \S\ref{results3d} for three-dimensional (3D) {\bf discs}, and we further
discuss them in \S\ref{discussions}. We summarise in \S\ref{summary} 
with a discussion on speculations for future work.

\input{model}
\input{results}
\input{results_3d}
\input{discussion}
\input{summary}

{\bf
\section*{Acknowledgments}
This material is based upon work supported by the National Science Foundation under Grant No. 1228509.
}
\bibliographystyle{mn2e}
\bibliography{ref}

\appendix
\input{appendix}

\end{document}

% In the case of protoplanetary discs, which have complex structure
% \citep{armitage10}, GI can also occur in `dead
% zones' --- sub-portions of the disc where the magneto-rotational
% instability is ineffective for angular momentum transport
% \citep{gammie96,turner08,landry13} and therefore has a reduced mass
% accretion rate. Dead zones may then become
% self-gravitating with sufficient mass built-up
% \citep{armitage01,martin12,martin12b,zhu09,zhu10,zhu10b,bae13,bae14}.

% original motivation for this work (DZGI)

%A wide range of GIs can occur in  gaseous 
%astrophysical discs 
%\citep{papaloizou89,christo92,christo93,hadley11,hadley14}.

% It is then
% of theoretical importance to clarify what type of GI is relevant to 
% partially self-gravitating discs, such as a protoplanetary disc containing a 
% massive dead zone. 
% This work presents a first step towards this goal. Specifically, here we 
% study a class of seemingly simple disc models which were found to 
% consistently produce \emph{one-armed} spiral structures, despite not 
% satisfying the usual requirements to generate them.

%through this work we identify a new method to generate one arm
%spirals 
%We therefore dedicate this study to verify their origin.

%brief history of m=1 
%stellar wobble from indiana group